---
title: "Stock prices data science project"
author: "Erika Harrell"
date: "2/3/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
### This project compares the high stock prices for 5 stocks: Apple, Netflix, CBS/Viacom, Amazon, and Disney. It took prices from the Yahoo Finance website <https://finance.yahoo.com/> for the dates from September 22, 2020 to December 22,2020.


```{r libraries}
#attach libraries
library(ggplot2)
library(ggthemes)
library(tidyr)
library(quantmod)
library(dplyr)
library(forecast)
```

## Loading Data
### Creating data folder, downloading datasets from Yahoo Finance and loading data into R Studio.
```{r load}
if(!dir.exists("./data")) {dir.create("./data")}
download.file("https://query1.finance.yahoo.com/v7/finance/download/AMZN?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/amazon.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/AAPL?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/apple.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/VIAC?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/cbs.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/DIS?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/disney.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/NFLX?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true", "./data/netflix.csv")
amazon<-read.csv("./data/amazon.csv")
apple<-read.csv("./data/apple.csv")
disney<-read.csv("./data/disney.csv")
cbs<-read.csv("./data/cbs.csv")
netflix<-read.csv("./data/netflix.csv")
```

## Data Wrangling
### Look at individual datasets, merge into a single dataset, and look at merged dataset.
```{r wrangle, echo=FALSE}
#Details of individual data sets
summary(amazon)
str(amazon)
summary(apple)
str(apple)
summary(cbs)
str(cbs)
summary(disney)
str(disney)
summary(netflix)
str(netflix)

#Rename variables in each dataset and combine into a single data set
data<-cbind(data.frame(Date=as.Date(amazon$Date),
            amazonOpen=round(as.numeric(amazon$Open),2),
            amazonHigh=round(as.numeric(amazon$High),2),
            amazonLow=round(as.numeric(amazon$Low),2),
            amazonClose=round(as.numeric(amazon$Close),2),
            amazonAdj.Close=round(as.numeric(amazon$Adj.Close),2),
            amazonVolume=round(as.numeric(amazon$Volume),0),
            appleOpen=round(as.numeric(apple$Open),2),
            appleHigh=round(as.numeric(apple$High),2),
            appleLow=round(as.numeric(apple$Low),2),
            appleClose=round(as.numeric(apple$Close),2),
            appleAdj.Close=round(as.numeric(apple$Adj.Close),2),
            appleVolume=round(as.numeric(apple$Volume),0),
            cbsOpen=round(as.numeric(cbs$Open),2),
            cbsHigh=round(as.numeric(cbs$High),2),
            cbsLow=round(as.numeric(cbs$Low),2),
            cbsClose=round(as.numeric(cbs$Close),2),
            cbsAdj.Close=round(as.numeric(cbs$Adj.Close),2),
            cbsVolume=round(as.numeric(cbs$Volume),0),
            disneyOpen=round(as.numeric(disney$Open),2),
            disneyHigh=round(as.numeric(disney$High),2),
            disneyLow=round(as.numeric(disney$Low),2),
            disneyClose=round(as.numeric(disney$Close),2),
            disneyAdj.Close=round(as.numeric(disney$Adj.Close),2),
            disneyVolume=round(as.numeric(disney$Volume),0),
            netflixOpen=round(as.numeric(netflix$Open),2),
            netflixHigh=round(as.numeric(netflix$High),2),
            netflixLow=round(as.numeric(netflix$Low),2),
            netflixClose=round(as.numeric(netflix$Close),2),
            netflixyAdj.Close=round(as.numeric(netflix$Adj.Close),2),
            netflixVolume=round(as.numeric(netflix$Volume),0)
            ))
#Details of combined dataset
summary(data)
str(data)
```

## Amazon forecasting analysis
### The line chart of the Amazon data with the associated regression line shows a a slight decrease in the high price over time. When the time series was decomposed, the trend showed a clear decrease.  The seasonal trend showed 2 period where there was a clear decrease in the high price. tEh plot of the forecasted data along with the observed dsata showed that the forecasted results did fall withint the 95% confidence level of the observed data. In terms of accuracy, the root mean squared erro was 61% for the training data and 65% for the testing data, showing that the model did not do well in predicting future high procies of the Amazon stock.
```{r amazon}
ggplot(data,aes(Date, amazonHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Amazon Sept 22-Dec 22")

#ts function creates time series object
ts1 <- ts(data$amazonHigh,frequency=32)
#plot time series object(daily high prices for Amazon)
plot(ts1,xlab="Days+1", ylab="AMZN")
plot(decompose(ts1),xlab="Days+1")

#training & test sets-have to build sets with consecutive time points
ts1Train <- window(ts1,start=1,end=1.7)
#window function creates test set that starts at time point 1.8
ts1Test <- window(ts1,start=1.8,end=2.97)
#show test data
ts1Train

#simple moving average
#plot training data 
#moving average (ma function) to plot)
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")

#exponential smoothing
#fit model that had different types of trends you want to fit
ets1 <- ets(ts1Train)
#get predictions and prediction bounds with forecast function
fcast <- forecast(ets1)
plot(fcast); 
lines(ts1Test,col="red")

#get accuracy
#accuracy(forecast,test set)
accuracy(fcast,ts1Test)
```

## Apple forecasting analysis
```{r apple}
ggplot(data,aes(Date, appleHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Apple Sept 22-Dec 22")

#ts function creates time series object
ts1 <- ts(data$appleHigh,frequency=32)
#plot time series object(daily high prices for Apple)
plot(ts1,xlab="Days+1", ylab="AAPL")
plot(decompose(ts1),xlab="Days+1")

#training & test sets-have to build sets with consecutive time points
ts1Train <- window(ts1,start=1,end=1.7)
#window function creates test set that starts at time point 1.8
ts1Test <- window(ts1,start=1.8,end=2.97)
#show test data
ts1Train

#simple moving average
#plot training data 
#moving average (ma function) to plot)
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")

#exponential smoothing
#fit model that had different types of trends you want to fit
ets1 <- ets(ts1Train)
#get predictions and prediction bounds with forecast function
fcast <- forecast(ets1)
plot(fcast); 
lines(ts1Test,col="red")

#get accuracy
#accuracy(forecast,test set)
accuracy(fcast,ts1Test)
```

## CBS/Viacom forecasting analysis
```{r cbs}
ggplot(data,aes(Date, cbsHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for CBS/Viacom Sept 22-Dec 22")

#ts function creates time series object
ts1 <- ts(data$cbsHigh,frequency=32)
#plot time series object(daily high prices for CBS/Viacom)
plot(ts1,xlab="Days+1", ylab="VIAC")
plot(decompose(ts1),xlab="Days+1")

#training & test sets-have to build sets with consecutive time points
ts1Train <- window(ts1,start=1,end=1.7)
#window function creates test set that starts at time point 1.8
ts1Test <- window(ts1,start=1.8,end=2.97)
#show test data
ts1Train

#simple moving average
#plot training data 
#moving average (ma function) to plot)
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")

#exponential smoothing
#fit model that had different types of trends you want to fit
ets1 <- ets(ts1Train)
#get predictions and prediction bounds with forecast function
fcast <- forecast(ets1)
plot(fcast); 
lines(ts1Test,col="red")

#get accuracy
#accuracy(forecast,test set)
accuracy(fcast,ts1Test)
```

## Disney forecasting analysis
```{r disney}
ggplot(data,aes(Date, disneyHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Disney Sept 22-Dec 22")

#ts function creates time series object
ts1 <- ts(data$disneyHigh,frequency=32)
#plot time series object(daily high prices for Disney)
plot(ts1,xlab="Days+1", ylab="DIS")
plot(decompose(ts1),xlab="Days+1")

#training & test sets-have to build sets with consecutive time points
ts1Train <- window(ts1,start=1,end=1.7)
#window function creates test set that starts at time point 1.8
ts1Test <- window(ts1,start=1.8,end=2.97)
#show test data
ts1Train

#simple moving average
#plot training data 
#moving average (ma function) to plot)
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")

#exponential smoothing
#fit model that had different types of trends you want to fit
ets1 <- ets(ts1Train)
#get predictions and prediction bounds with forecast function
fcast <- forecast(ets1)
plot(fcast); 
lines(ts1Test,col="red")

#get accuracy
#accuracy(forecast,test set)
accuracy(fcast,ts1Test)
```

## Netflix forecasting analysis
```{r netflix}
ggplot(data,aes(Date, netflixHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Netflix Sept 22-Dec 22")

#ts function creates time series object
ts1 <- ts(data$netflixHigh,frequency=32)
#plot time series object(daily high prices for Netflix)
plot(ts1,xlab="Days+1", ylab="NFLX")
plot(decompose(ts1),xlab="Days+1")

#training & test sets-have to build sets with consecutive time points
ts1Train <- window(ts1,start=1,end=1.7)
#window function creates test set that starts at time point 1.8
ts1Test <- window(ts1,start=1.8,end=2.97)
#show test data
ts1Train

#simple moving average
#plot training data 
#moving average (ma function) to plot)
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")

#exponential smoothing
#fit model that had different types of trends you want to fit
ets1 <- ets(ts1Train)
#get predictions and prediction bounds with forecast function
fcast <- forecast(ets1)
plot(fcast); 
lines(ts1Test,col="red")

#get accuracy
#accuracy(forecast,test set)
accuracy(fcast,ts1Test)
```

