---
title: "Stock prices data science project"
author: "Erika Harrell"
date: "1/8/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
### This project compars stock prices for 5 stocks: Apple, Netflix, CBS/Viacom, Amazon, and Disney. It took pirces from the Yahoo Finance website <https://finance.yahoo.com/> for the dates from September 22, 2020 to December 22,2020.


```{r libraries}
#adding libraries
library(ggplot2)
library(ggthemes)
library(tidyr)
library(dplyr)
```

## Loading Data-creating data folder, downloading datasets from Yahoo Finance and loading data into R Studio.
```{r load, echo=FALSE}
if(!dir.exists("./data")) {dir.create("./data")}
download.file("https://query1.finance.yahoo.com/v7/finance/download/AMZN?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/amazon.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/AAPL?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/apple.csv")
download.file("https://query1.finance.yahoo.com/v7/finance/download/VIAC?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/cbs.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/DIS?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true","./data/disney.csv")

download.file("https://query1.finance.yahoo.com/v7/finance/download/NFLX?period1=1600819200&period2=1608681600&interval=1d&events=history&includeAdjustedClose=true", "./data/netflix.csv")
amazon<-read.csv("./data/amazon.csv")
apple<-read.csv("./data/apple.csv")
disney<-read.csv("./data/disney.csv")
cbs<-read.csv("./data/cbs.csv")
netflix<-read.csv("./data/netflix.csv")
```

## Data Wrangling
### Look at individual datasets, merge into a single dataset, and look at single dataset.
```{r wrangle, echo=FALSE}
#Details of individual data sets
summary(amazon)
str(amazon)
summary(apple)
str(apple)
summary(cbs)
str(cbs)
summary(disney)
str(disney)
summary(netflix)
str(netflix)

#Rename variables in each dataset and combine into a single data set
data<-cbind(data.frame(Date=as.Date(amazon$Date),
            amazonOpen=round(as.numeric(amazon$Open),2),
            amazonHigh=round(as.numeric(amazon$High),2),
            amazonLow=round(as.numeric(amazon$Low),2),
            amazonClose=round(as.numeric(amazon$Close),2),
            amazonAdj.Close=round(as.numeric(amazon$Adj.Close),2),
            amazonVolume=round(as.numeric(amazon$Volume),0),
            appleOpen=round(as.numeric(apple$Open),2),
            appleHigh=round(as.numeric(apple$High),2),
            appleLow=round(as.numeric(apple$Low),2),
            appleClose=round(as.numeric(apple$Close),2),
            appleAdj.Close=round(as.numeric(apple$Adj.Close),2),
            appleVolume=round(as.numeric(apple$Volume),0),
            cbsOpen=round(as.numeric(cbs$Open),2),
            cbsHigh=round(as.numeric(cbs$High),2),
            cbsLow=round(as.numeric(cbs$Low),2),
            cbsClose=round(as.numeric(cbs$Close),2),
            cbsAdj.Close=round(as.numeric(cbs$Adj.Close),2),
            cbsVolume=round(as.numeric(cbs$Volume),0),
            disneyOpen=round(as.numeric(disney$Open),2),
            disneyHigh=round(as.numeric(disney$High),2),
            disneyLow=round(as.numeric(disney$Low),2),
            disneyClose=round(as.numeric(disney$Close),2),
            disneyAdj.Close=round(as.numeric(disney$Adj.Close),2),
            disneyVolume=round(as.numeric(disney$Volume),0),
            netflixOpen=round(as.numeric(netflix$Open),2),
            netflixHigh=round(as.numeric(netflix$High),2),
            netflixLow=round(as.numeric(netflix$Low),2),
            netflixClose=round(as.numeric(netflix$Close),2),
            netflixyAdj.Close=round(as.numeric(netflix$Adj.Close),2),
            netflixVolume=round(as.numeric(netflix$Volume),0)
            ))
#Details of combined dataset
summary(data)
str(data)
```

## Plots of High prices

```{r plots, echo=FALSE}
ggplot(data,aes(Date, amazonHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Amazon Sept 22-Dec 22")
ggplot(data,aes(Date, appleHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Apple Sept 22-Dec 22")
ggplot(data,aes(Date, cbsHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for CBS Sept 22-Dec 22")
ggplot(data,aes(Date, disneyHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Disney Sept 22-Dec 22")
ggplot(data,aes(Date, netflixHigh, group=1))+
  geom_line(color="darkblue" ,size=2)+
  geom_smooth(method=lm,se=FALSE, colour="black")+
  ggtitle("High stock prices for Netflix Sept 22-Dec 22")
par(mfrow=c(1,1))
```
